### 2024-11-08 Article/Background_sections/Adhesivity.tex
some proofs
last proofs

\begin{lemma}
    Let $\cat C$ be an $\mathcal{M}$-adhesive category with all pullbacks, and suppose that the cube below is given, in which every face is a pullback and the bottom one is a $\mathcal{M}$-pushout.
        \[\begin{tikzcd}[row sep=25, column sep=25]
	& {A'} && {B'} \\
	{C'} && {D'} \\
	& A && B \\
	C && D
	\arrow["{f'}", from=1-2, to=1-4]
	\arrow["{m'}"', from=1-2, to=2-1]
	\arrow["a"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{n'}"', from=1-4, to=2-3]
	\arrow["b", from=1-4, to=3-4]
	\arrow["{g'}"{pos=0.7}, from=2-1, to=2-3, crossing over]
	\arrow["c"', from=2-1, to=4-1]
	\arrow["f"'{pos=0.3}, from=3-2, to=3-4]
	\arrow["d"{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["m", from=3-2, to=4-1]
	\arrow["n", from=3-4, to=4-3]
	\arrow["g"', from=4-1, to=4-3]
    \end{tikzcd}\]

@@

    {\color{red} Non ho capito la dimostrazione!!}
\end{proof}

>>>>

\begin{lemma}\label{lemma:pushouts_kernel_pairs}
    Let $\cat C$ be an $\mathcal{M}$-adhesive category with all pullbacks, and suppose that the cube below is given, in which every face is a pullback and the bottom one is a $\mathcal{M}$-pushout.
        \[\begin{tikzcd}[row sep=25, column sep=25]
	& {A'} && {B'} \\
	{C'} && {D'} \\
	& A && B \\
	C && D
	\arrow["{f'}", from=1-2, to=1-4]
	\arrow["{m'}"', from=1-2, to=2-1]
	\arrow["a"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{n'}"', from=1-4, to=2-3]
	\arrow["b", from=1-4, to=3-4]
	\arrow["{g'}"{pos=0.7}, from=2-1, to=2-3, crossing over]
	\arrow["c"', from=2-1, to=4-1]
	\arrow["f"'{pos=0.3}, from=3-2, to=3-4]
	\arrow["d"{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["m", from=3-2, to=4-1]
	\arrow["n", from=3-4, to=4-3]
	\arrow["g"', from=4-1, to=4-3]
    \end{tikzcd}\]

@@

	$f'$ is in $\mathcal{M}$ as it is the pullback of $\mathcal{M}$, thus the bottom face of the cube is a Van Kampen pushout and the thesis follows.

### 2024-11-08 Article/Graph_sections/EGGs.tex
last proofs

@@

@@

\begin{proof}
	TODO	
\end{proof}

>>>>

\iffalse

@@

\fi

E-Graphs are a particular type of graphs with equivalences.

\begin{definition}[E-Graph]\label{def:e_graph}
	Given a graph with equivalence $\eqgraph{G} = (E, V, C, s, t, q)$, let $(S, p_1, p_2)$ be the kernel pair of $q \circ s$. Then, $\eqgraph G$ is an \emph{e-graph} if it holds that
	\[
		q \circ t \circ p_1 = q \circ t \circ p_2
	\]
\end{definition}

\begin{remark}
	From a set-theoretic point of view, $\eqgraph{G} = (\graph{G}, \sim)$ is a graph with equivalence, where $\graph{G} = (E, V, s, t)$ in which holds that, for each pair of edges $e$, $e'$,
	\[
		\frac{s(e)\sim s(e')}{t(e)\sim t(e')}
	\]
	As we have considered in the section on graphs with equivalences, we can formalize the equivalence relation with a surjective function $q: V \to V/_\sim = C$, rewriting the inference rule above as an inclusion.
	Let $S = \{(e, e') \in E\times E \mid q(s(e)) = q(s(e'))\}$ and $T = \{(e, e') \in E \times E \mid q(t(e)) = q(t(e'))\}$. Then, $\eqgraph{G}$ is an e-graph if $S \subseteq T$. As we noted in \Cref{ex:kernel_pairs_in_Set}, $S$ is the kernel pair of $q \circ s$ and $T$ is the kernel pair of $q \circ t$, if endowed with canonical projections. Moreover, we notice that the pairing of such projections yields a subobject of the product $E \times E$ \Cref{prop:pairng_of_kernel_pairs_mono}. Hence, by \Cref{rem:fact_of_subobject_is_unique}, the \Cref{def:e_graph} express this fact using categorial constructions.
\end{remark}

\begin{definition}[Category of E-Graphs]\label{def:cat_of_eggs}
	$\EGG$ is the subcategory of $\EqGrph$ in which objects are e-graphs, and, given two e-graphs $\eqgraph{G, H}$, $\EGG(\eqgraph{G, H}) = \EqGrph(\eqgraph{G, H})$.
\end{definition}

@@

\begin{proof}
	Let $\eqgraph{G}_1 = (A_1, B_1, C_1, s_1, t_1, q_1)$, $\eqgraph{G}_2 = (A_2, B_2, C_2, s_2, t_2, q_2)$ and $\eqgraph{G}_3 = (A_3, B_3, C_3, s_3, t_3, q_3)$ be e-graphs,  $h: \eqgraph{G}_1 \to \eqgraph{G}_2$ and $m: \eqgraph{G}_1 \to \eqgraph{G}_3$ be two morphisms with $m$ mono. Let now $\eqgraph{P} = (A, B, C, s, t, q)$, together with $k: \eqgraph{G}_3 \to \eqgraph{P}$ and $n: \eqgraph{G}_2 \to \eqgraph{P}$, the pushout of $h$ and $m$.
	Let $(K_i, \pi_i^1, \pi_i^2)$ the kernel pair of $q_i\circ s_i$, for $i = 1, 2, 3$, and let $(U, u_1, u_2)$ be the kernel pair of $q \circ s$.
	Consider the following cube in $\Set$
	\[\begin{tikzcd}[row sep=25, column sep = 25]
	& {A_1} && {A_2} \\
	{A_3} && A \\
	& {C_1} && {C_2} \\
	{C_3} && C
	\arrow["{h_E}", from=1-2, to=1-4]
	\arrow["{m_E}"', from=1-2, to=2-1]
	\arrow["{q_1\circ s_1}"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{n_E}", from=1-4, to=2-3]
	\arrow["{q_2\circ s_2}", from=1-4, to=3-4]
	\arrow["{k_E}"{pos=0.7}, from=2-1, to=2-3, crossing over]
	\arrow["{q_3\circ s_3}"', from=2-1, to=4-1]
	\arrow["{h_C}"{pos=0.8}, from=3-2, to=3-4]
	\arrow["{m_C}"', from=3-2, to=4-1]
	\arrow["{n_C}", from=3-4, to=4-3]
	\arrow["{k_C}"', from=4-1, to=4-3]
	\arrow["{q\circ s}"'{pos=0.2}, from=2-3, to=4-3, crossing over]
	\end{tikzcd}\]
	By adhesivity of $\Set$, and since $m$ is regular mono, the cube above is Van Kampen, and, by \Cref{prop:monos_are_preserved_by_pullbacks_in_adh_cats}, all the faces are pullbacks. Then, applying \Cref{lemma:pushouts_kernel_pairs}, the square below is a pushout
	\[\begin{tikzcd}[row sep=27, column sep = 27]
		K_1 \ar[r, "{f_h}"] \ar[d, "{f_m}"swap] & K_2 \ar[d, "{f_n}"] \\
		K_3 \ar[r, "{f_k}"swap] & U
	\end{tikzcd}\]

	Hence, we end up with the following situation
	\[
		\begin{tikzcd}[row sep = 20, column sep = 20]
	& {K_1} && {K_2} \\
	{K_3} && U \\
	& {A_1} && {A_2} \\
	{A_3} && A \\
	& {C_1} && {C_2} \\
	{C_3} && C
	\arrow["{f_h}", from=1-2, to=1-4]
	\arrow["{f_m}"', from=1-2, to=2-1]
	\arrow["{\pi_1^1}"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{f_n}", from=1-4, to=2-3]
	\arrow["{\pi^1_2}", from=1-4, to=3-4]
	\arrow["{f_k}"{pos=0.8}, from=2-1, to=2-3, crossing over]
	\arrow["{\pi_3^1}"', from=2-1, to=4-1]
	\arrow["{u_1}"'{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["{h_E}"{pos=0.8}, from=3-2, to=3-4]
	\arrow["{m_E}"', from=3-2, to=4-1]
	\arrow["{q_1\circ s_1}"'{pos=0.7}, from=3-2, to=5-2]
	\arrow["{n_E}", from=3-4, to=4-3]
	\arrow["{q_2\circ s_2}", from=3-4, to=5-4]
	\arrow["{k_E}"{pos=0.7}, from=4-1, to=4-3, crossing over]
	\arrow["{q_3\circ s_3}"', from=4-1, to=6-1]
	\arrow["{h_C}"{pos=0.8}, from=5-2, to=5-4]
	\arrow["{q\circ s}"'{pos=0.2}, from=4-3, to=6-3, crossing over]
	\arrow["{m_C}"', from=5-2, to=6-1]
	\arrow["{n_C}", from=5-4, to=6-3]
	\arrow["{k_C}"', from=6-1, to=6-3]
	\end{tikzcd}
		\qquad
		\begin{tikzcd}[row sep = 20, column sep = 20]
	& {K_1} && {K_2} \\
	{K_3} && U \\
	& {A_1} && {A_2} \\
	{A_3} && A \\
	& {C_1} && {C_2} \\
	{C_3} && C
	\arrow["{f_h}", from=1-2, to=1-4]
	\arrow["{f_m}"', from=1-2, to=2-1]
	\arrow["{\pi_1^2}"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{f_n}", from=1-4, to=2-3]
	\arrow["{\pi^2_2}", from=1-4, to=3-4]
	\arrow["{f_k}"{pos=0.8}, from=2-1, to=2-3, crossing over]
	\arrow["{\pi_3^2}"', from=2-1, to=4-1]
	\arrow["{u_2}"'{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["{h_E}"{pos=0.8}, from=3-2, to=3-4]
	\arrow["{m_E}"', from=3-2, to=4-1]
	\arrow["{q_1\circ s_1}"'{pos=0.7}, from=3-2, to=5-2]
	\arrow["{n_E}", from=3-4, to=4-3]
	\arrow["{q_2\circ s_2}", from=3-4, to=5-4]
	\arrow["{k_E}"{pos=0.7}, from=4-1, to=4-3, crossing over]
	\arrow["{q_3\circ s_3}"', from=4-1, to=6-1]
	\arrow["{h_C}"{pos=0.8}, from=5-2, to=5-4]
	\arrow["{q\circ s}"'{pos=0.2}, from=4-3, to=6-3, crossing over]
	\arrow["{m_C}"', from=5-2, to=6-1]
	\arrow["{n_C}", from=5-4, to=6-3]
	\arrow["{k_C}"', from=6-1, to=6-3]
		\end{tikzcd}\]

	Computing, we have
	\[
		\begin{split}
			q \circ t \circ u_1 \circ f_n &= q \circ t \circ n_E \circ \pi_2^1 \\ &= n_C \circ q_2 \circ s_2 \circ \pi_2^1 \\&= n_C \circ q_2 \circ s_2 \circ \pi_2^2 \\&=q \circ t \circ u_2 \circ f_n 
			\end{split}
			\qquad
			\begin{split} q \circ t \circ u_1 \circ f_k &= q \circ t \circ k_E \circ \pi_3^1 \\ &= k_C \circ q_3 \circ s_3 \circ \pi_3^1 \\&= k_C \circ q_3 \circ s_3 \circ \pi_3^2 \\&=q \circ t \circ u_2 \circ f_k
			\end{split}
	\]
	By universal property of pushouts, we can conclude $q \circ t \circ u_1 = q \circ t \circ u_2$, hence the thesis.
\end{proof}

\begin{cor}
	$I$ creates pushouts along regular monos.
\end{cor}

By direct application of \Cref{th:crit_for_adh}, we can conclude what follows.{rem:fact_of_subobject_is_unique}

\begin{cor}
	$\EGG$ is $\Reg(\EGG)-adhesive$.
\end{cor}
